Long-term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-08-29 article/article.Rmd
Commit 2024-08-29 20:14

<!-- extreme caces criterion -->
A sub-category of the CE events that is often discussed in the literature
[@Cordero2023; @Martins2022; @Yordanov2015], are the extreme cloud enhancement (ECE)
events. These are cases of CE where the measured intensity of the irradiance exceeds
the TSI at the top of the atmosphere. In this case the threshold $E$ is given by:
<!--and detected according to the Equation\nobreakspace{}\ref{eq:ECE}.-->
\begin{equation}
\text{ECE}: E > \cos(\text{SZA}) \cdot E_{\odot} \frac{r^2_\text{m}} {r^2}
\label{eq:ECE}
\end{equation}
where: $\text{SZA}$ the solar zenith angle, $E_{\odot}$ the solar constant adjusted for
the actual Sun -- Earth distance $r$ and $r_\text{m}$ is the mean Sun -- Earth
distance of $1.496\times10^8\,\text{km}$.

@@

Figure\nobreakspace{}\ref{fig:P-energy} shows the time series of the yearly number of
CE cases (each with duration of one minute), the yearly mean OI and the yearly
excess irradiation for the period 1993 -- 2023, together with corresponding linear
trends. All three quantities show increasing trends, most pronounced for the
frequency of occurrence 
($`r plus_sign(round(yearlyCE[var == "GLB_diff.N", slope], 1))`\pm `r round(2 * yearlyCE[var == "GLB_diff.N", slope.sd], 1)`\,\text{cases}/\text{year}$)
and the excess irradiation 
($`r plus_sign(round(yearlyCE[var == "GLB_diff.sum", slope], 1))` \pm `r round(2 * yearlyCE[var == "GLB_diff.sum", slope.sd], 1)`\,\text{kJ}/\text{year}$),
which are also statistically significant.
In contrast the trend of the yearly mean OI is negligible 
($`r plus_sign(round(yearlyCE[var == "GLB_diff.mean", slope], 1))`\pm `r round(2 * yearlyCE[var == "GLB_diff.mean", slope.sd], 1)`\,\text{W}/\text{m}^2/\text{year}$)
and of no statistical significance.
The average OI for the entire period is
$`r plus_sign(round(mean(ST_E_yearly$GLB_diff.mean), 1))`\pm `r round(sd(ST_E_yearly$GLB_diff.mean),1)`\,\text{W}/\text{m}^2$.
The interannual variability of the data about the trend lines is quite large.
Furthermore, the spread tends to increase with time (at least for the quantities of
panels b and c), suggesting a significant variability in cloud patterns over the
area, possibly associated to changes in climate.

@@

```{r P-extreme-distribution, echo=FALSE, fig.cap="Distribution of ECE events in Thessaloniki for the period 1993 -- 2023."}
knitr::include_graphics("../images/P-extreme-distribution-1.png")
```

>>>>

<!-- extreme caces criterion -->
A sub-category of the CE events that is often discussed in the literature
[@Cordero2023; @Martins2022; @Yordanov2015], are the extreme cloud enhancement (ECE)
events. These are cases of CE where the measured intensity of the irradiance exceeds
the TSI at the top of the atmosphere. In this case the threshold $E$ is given by:
<!--and detected according to the Equation\nobreakspace{}\ref{eq:ECE}.-->
\begin{equation}
\text{ECE}: E > \cos(\text{SZA}) \cdot E_{\odot} \frac{r^2_\text{m}} {r^2}
\label{eq:ECE}
\end{equation}
where: $\text{SZA}$ the solar zenith angle, $E_{\odot}$ is the solar constant at the
mean Sun - Earth distance ($`r round(solar_constant)`\,\text{W}/\text{m}^2$), $r$ is
the actual Sun - Earth distance and $r_\text{m}$ is the mean Sun -- Earth distance of
$1.496\times10^8\,\text{km}$.

@@

Figure\nobreakspace{}\ref{fig:P-energy} shows the time series of the yearly number of
CE cases (each with duration of one minute), the yearly mean OI and the yearly
excess irradiation for the period 1993 -- 2023, together with corresponding linear
trends. All three quantities show increasing trends, most pronounced for the
frequency of occurrence 
($`r plus_sign(round(yearlyCE[var == "GLB_diff.N", slope], 1))`\pm `r round(2 * yearlyCE[var == "GLB_diff.N", slope.sd], 1)`\,\text{cases}/\text{year}$)
and the excess irradiation 
($`r plus_sign(round(yearlyCE[var == "GLB_diff.sum", slope], 1))` \pm `r round(2 * yearlyCE[var == "GLB_diff.sum", slope.sd], 1)`\,\text{kJ}/\text{year}$),
which are also statistically significant.
In contrast the trend of the yearly mean OI is negligible 
($`r plus_sign(round(yearlyCE[var == "GLB_diff.mean", slope], 2))`\pm `r round(2 * yearlyCE[var == "GLB_diff.mean", slope.sd], 1)`\,\text{W}/\text{m}^2/\text{year}$)
and of no statistical significance.
The average OI for the entire period is
$`r plus_sign(round(mean(ST_E_yearly$GLB_diff.mean), 1))`\pm `r round(sd(ST_E_yearly$GLB_diff.mean),1)`\,\text{W}/\text{m}^2$.
The interannual variability of the data about the trend lines is quite large.
Furthermore, the spread tends to increase with time (at least for the quantities of
panels b and c), suggesting a significant variability in cloud patterns over the
area, possibly associated to changes in climate.

@@

```{r P-extreme-distribution, echo=FALSE, fig.cap="Distribution of ECE events in Thessaloniki for the period 1993 -- 2023. Measured as extra irradiance over TSI at TOA for the same SZA."}
knitr::include_graphics("../images/P-extreme-distribution-2.png")
```

<!-- ```{r P-extreme-distribution, echo=FALSE, fig.cap="Distribution of ECE events in Thessaloniki for the period 1993 -- 2023."} -->
<!-- knitr::include_graphics("../images/P-extreme-distribution-1.png") -->
<!-- ``` -->

### 2024-08-29 article/article.tex
Commit 2024-08-29 20:14

A sub-category of the CE events that is often discussed in the
literature \citep{Cordero2023, Martins2022, Yordanov2015}, are the
extreme cloud enhancement (ECE) events. These are cases of CE where the
measured intensity of the irradiance exceeds the TSI at the top of the
atmosphere. In this case the threshold \(E\) is given by:
\begin{equation}
\text{ECE}: E > \cos(\text{SZA}) \cdot E_{\odot} \frac{r^2_\text{m}} {r^2}
\label{eq:ECE}
\end{equation} where: \(\text{SZA}\) the solar zenith angle,
\(E_{\odot}\) the solar constant adjusted for the actual Sun -- Earth
distance \(r\) and \(r_\text{m}\) is the mean Sun -- Earth distance of
\(1.496\times10^8\,\text{km}\).

@@

Figure\nobreakspace{}\ref{fig:P-energy} shows the time series of the
yearly number of CE cases (each with duration of one minute), the yearly
mean OI and the yearly excess irradiation for the period 1993 -- 2023,
together with corresponding linear trends. All three quantities show
increasing trends, most pronounced for the frequency of occurrence
(\(+45.6\pm 22\,\text{cases}/\text{year}\)) and the excess irradiation
(\(+116.3 \pm 68.2\,\text{kJ}/\text{year}\)), which are also
statistically significant. In contrast the trend of the yearly mean OI
is negligible (\(+0\pm 0.1\,\text{W}/\text{m}^2/\text{year}\)) and of no
statistical significance. The average OI for the entire period is
\(+40.1\pm 2.5\,\text{W}/\text{m}^2\). The interannual variability of
the data about the trend lines is quite large. Furthermore, the spread
tends to increase with time (at least for the quantities of panels b and
c), suggesting a significant variability in cloud patterns over the
area, possibly associated to changes in climate.

@@

{\centering \includegraphics[width=1\linewidth]{../images/P-extreme-distribution-1}

@@

\caption{Distribution of ECE events in Thessaloniki for the period 1993 -- 2023.}\label{fig:P-extreme-distribution}
\end{figure}

>>>>

A sub-category of the CE events that is often discussed in the
literature \citep{Cordero2023, Martins2022, Yordanov2015}, are the
extreme cloud enhancement (ECE) events. These are cases of CE where the
measured intensity of the irradiance exceeds the TSI at the top of the
atmosphere. In this case the threshold \(E\) is given by:
\begin{equation}
\text{ECE}: E > \cos(\text{SZA}) \cdot E_{\odot} \frac{r^2_\text{m}} {r^2}
\label{eq:ECE}
\end{equation} where: \(\text{SZA}\) the solar zenith angle,
\(E_{\odot}\) is the solar constant at the mean Sun - Earth distance
(\(1361\,\text{W}/\text{m}^2\)), \(r\) is the actual Sun - Earth
distance and \(r_\text{m}\) is the mean Sun -- Earth distance of
\(1.496\times10^8\,\text{km}\).

@@

Figure\nobreakspace{}\ref{fig:P-energy} shows the time series of the
yearly number of CE cases (each with duration of one minute), the yearly
mean OI and the yearly excess irradiation for the period 1993 -- 2023,
together with corresponding linear trends. All three quantities show
increasing trends, most pronounced for the frequency of occurrence
(\(+45.6\pm 22\,\text{cases}/\text{year}\)) and the excess irradiation
(\(+116.3 \pm 68.2\,\text{kJ}/\text{year}\)), which are also
statistically significant. In contrast the trend of the yearly mean OI
is negligible (\(+0.03\pm 0.1\,\text{W}/\text{m}^2/\text{year}\)) and of
no statistical significance. The average OI for the entire period is
\(+40.1\pm 2.5\,\text{W}/\text{m}^2\). The interannual variability of
the data about the trend lines is quite large. Furthermore, the spread
tends to increase with time (at least for the quantities of panels b and
c), suggesting a significant variability in cloud patterns over the
area, possibly associated to changes in climate.

@@

{\centering \includegraphics[width=1\linewidth]{../images/P-extreme-distribution-2}

@@

\caption{Distribution of ECE events in Thessaloniki for the period 1993 -- 2023. Measured as extra irradiance over TSI at TOA for the same SZA.}\label{fig:P-extreme-distribution}
\end{figure}
